Finding a closest point in a lattice of Voronoi's first kind

### 2014-07-09 reviews/response.tex

The formula for $\|B\xbf\|_2^2$ stated by the reviewer is known as~\emph{Selling's formula} in literature on quadratic forms and the geometry of numbers~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  We do indirectly use this formula in the form of the function $\Phi(S,p)$ and our Lemma~\zref{lem:decSellings}.  A very early draft of our paper attempted to use Selling's formula more directly, but we ultimately found that it was $\Phi(S,p)$ that was used repeatedly and not Selling's formula.

\item\textbf{Comment:} 
To present the result from the perspective of Computational Complexity (also not given in the
paper), this paper can be viewed as giving a polynomial time algorithm for the Closest Vector
Problem with Preprocessing (CVPP) over a special class of lattices. In the preprocessing model,
we are allowed to compute any polynomial amount of advice about the lattice (i.e. a good basis, or many short vectors in the lattice, etc.) before answering any CVP queries, where the resources needed to compute this advice are not counted in the runtime of the algorithm. In its
gap decisional version, where we need only decide whether a target is at distance $\leq d$ or $> \alpha d$
(for some approximation factor $\alpha \geq 1$), it is was shown [1, 4] to be NP-Hard for any constant $\alpha = 2^{\log^{1-\epsilon}n}$, and to not be polynomial time solvable for $\alpha = 2 \log n$ , for any fixed $\epsilon > 0$, under the assumption  that $NP \subsetneq \text{RTIME}(2^{\log^{O(1/\epsilon)}n})$ (quasi-polynomial time). For its approximate search version (i.e. find a lattice vector whose distance to the target is some bounded factor away from optimal), an $O(n^{1.5})$ approximation algorithm was (implicitly) given in [5], which was only recently improved to $O( \sqrt{ n / \log n} )$ in [3].
\\\textbf{Response:}

\item\textbf{Comment:} 
Hence, from the standpoint of CVPP, the algorithm in this paper gives essentially the first
exact polynomial time algorithm over a reasonably large and natural class of lattices, i.e. those
of Voronoi’s first kind. Lastly, removing the need for preprocessing for these lattices, at least for
solving SVP without being given an obtuse superbasis, would require a substantial breakthrough
in the theory of lattice problems (not stated in paper). Indeed, by a simple reduction this would
allow us to decide whether a lattice (given any of its bases) is a rotation of $\ints^n$ (which is clearly
of Voronoi's first kind). This is a long standing open problem in the theory of lattices, which has
only recently been solved in certain very special cases~[6].
\\\textbf{Response:}
The final paragraph of Section~\zref{sec:discussion} has been modified to make the connection with the existing problem of determining whether a rectangular basis exists.

``A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.  It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.  Lattices of dimension larger than 3 need not be of Voronoi's first kind.  An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?  A related question is: is it possible to efficiently find a obtuse superbasis if it exists?  It is suspected that the answer to both of these questions is no.  An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.  It is suspected that this simpler problem is difficult.''

@@

\item\textbf{Comment:} 
This paper makes a solid contribution to the theory of lattice problems. I think the results here are
well-motivated (I encourage the authors to include some of the motivation mentioned in the summary), and point to the possibility that there might be large and useful classes of lattices where
solving CVP is “easy”.
\\\textbf{Response:}
The following paragraph has been added to Section~\zref{sec:discussion} to suggest that similar fast algorithms for other classes of lattices might exist:

``Another interesting question is: are there subfamilies of Voronoi's first kind that admit even faster algorithms?  Both $A_n$ and $A_n^*$ are examples of this, but there might exist other subfamilies with algorithms faster than $O(n^4)$.  A related question is: can the techniques developed in this paper be applied to other families of lattices, i.e., beyond just those of Voronoi's first kind?''

\item\textbf{Comment:} 
On a technical level, I found the main convergence proof to be interesting, as I've rarely seen
combinatorial arguments successfully applied to lattice problems. On the other hand, while I’ve
convinced myself the authors have the ``right'' convergence proof, I found its presentation to be
rather awkward and unintuitive. I highly recommend that the authors rephrase the proof using
the language of Laplacians, which would make things more clear (see the additional comments
below for some technical suggestions).
\\\textbf{Response:}
We thank the reviewer for their comments.  We respond to the authors concerns below.

@@

We now respond to the reviewers numbered comments.

\begin{enumerate}

\item\textbf{Comment:} 
I believe Proposition 2.1 and 2.2 appear essentially verbatim in [8]. Same for Propositions
3.1 (which is essentially by definition) and 3.2, they appear in [9]. I don’t think it adds any
readibility/anything of value to reprove them here.
\\\textbf{Response:}

We agree that these propositions are well known/straightforward, but would prefer to keep them so that the paper is self contained. We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).  Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.  As far as we are aware the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume) as in Proposition~2.2.  Instead, the proceedure operates on scaled versions of the lattice.  The scaling is chosen to ensure that the target point always lies inside twice the Voronoi cell.

A motivation for Propositions 3.1 and 3.2 is as follows.  The stopping criterion in our proceedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.  The stopping criterions in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013} essentially involve a direct check that a given point, say $x$, is contained inside the Voronoi cell.  That is, one directly asserts that $2 x \cdot v \leq \|v\|^2$ for all relevant vectors $v$.  We cannot use this approach since it would amount to checking this inequality over as many as $2^{n+1}-2$ relevant vectors.  Instead we simply check that $\|y - x_K\| \leq \|y - x_{K+1}\|$.  That is, we check that the proceedure converges in the sense that 
\[
\|y - x_K\| = \|y - x_{K+1}\| = \|y - x_{K+2}\| = \dots.
\]
This is a straightforward, but important, difference and it motivates Propostions 3.1 and 3.2.

### 2014-07-10 reviews/response.tex

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

\maketitle

We thank both reviewers for their thoughtful comments.  Below, the citations within the reviewer comments are left as given by the reviewer.  Citations in our responses refer to the references at the end of this document.

@@

\begin{itemize}

\item\textbf{Comment:} 
From the perspective of quadratic forms, the quadratic forms induced by obtuse superbases are in one to one correspondance with graph Laplacians (this is not
explicitly stated in the paper for some reason). Indeed, the quadratic form $\|B\xbf\|^2$ is exactly $\xbf^T L_G \xbf$, where $L_G$ is the Laplacian of the weighted graph $G$ on vertices $\{ 1, . . . , n + 1 \}$ , with edges weights $w_{ij} = -\bbf_i \cdot \bbf_j \geq 0, i = j$ (which they call the Selling parameters). More precisely:
\[
\|B\xbf\|_2^2 = \xbf^T L_G \xbf = \sum_{ \{i,j\} \in E[G] } w_{ij} (\xbf_{i} - \xbf_{j} )^2.
\]
\\ \textbf{Response:} We thank the reviewer for pointing out the connection with the Laplacian matrix of a graph.  The following text has been included in Section~\zref{sec:comp-clos-relev} to highlight this link.

``At the core of this technique is the fact that a one-to-one correspondence exists between the obtuse superbasis of a lattice of Voronoi's first kind and the~\emph{Laplacian matrix}~\cite{Chung_spectral_graph_theory_1997,Cvetković_spectra_graphs_1998} of a simple weighted graph with $n+1$ vertices and positive edge weights equal to the negated Selling parameters $-q_{ij}$.''

The formula for $\|B\xbf\|_2^2$ stated by the reviewer is known as~\emph{Selling's formula} in the literature on quadratic forms and the geometry of numbers~\cite[Proposition 2.3.1]{Valentin2003_coverings_tilings_low_dimension}~\cite{Selling1874}.  We do indirectly use this formula in the form of the function $\Phi(S,p)$ and our Lemma~\zref{lem:decSellings}.  A very early draft of our paper attempted to use Selling's formula more directly, but we ultimately found that it was $\Phi(S,p)$ that was used repeatedly and not Selling's formula.

@@

\item\textbf{Comment:} 
To present the result from the perspective of Computational Complexity (also not given in the
paper), this paper can be viewed as giving a polynomial time algorithm for the Closest Vector
Problem with Preprocessing (CVPP) over a special class of lattices. In the preprocessing model,
we are allowed to compute any polynomial amount of advice about the lattice (i.e. a good basis, or many short vectors in the lattice, etc.) before answering any CVP queries, where the resources needed to compute this advice are not counted in the runtime of the algorithm. In its
gap decisional version, where we need only decide whether a target is at distance $\leq d$ or $> \alpha d$
(for some approximation factor $\alpha \geq 1$), it is was shown [1, 4] to be NP-Hard for any constant $\alpha = 2^{\log^{1-\epsilon}n}$, and to not be polynomial time solvable for $\alpha = 2 \log n$ , for any fixed $\epsilon > 0$, under the assumption  that $NP \subsetneq \text{RTIME}(2^{\log^{O(1/\epsilon)}n})$ (quasi-polynomial time). For its approximate search version (i.e. find a lattice vector whose distance to the target is some bounded factor away from optimal), an $O(n^{1.5})$ approximation algorithm was (implicitly) given in [5], which was only recently improved to $O( \sqrt{ n / \log n} )$ in [3].
\\\textbf{Response:}

\item\textbf{Comment:} 
Hence, from the standpoint of CVPP, the algorithm in this paper gives essentially the first
exact polynomial time algorithm over a reasonably large and natural class of lattices, i.e. those
of Voronoi's first kind. Lastly, removing the need for preprocessing for these lattices, at least for
solving SVP without being given an obtuse superbasis, would require a substantial breakthrough
in the theory of lattice problems (not stated in paper). Indeed, by a simple reduction this would
allow us to decide whether a lattice (given any of its bases) is a rotation of $\ints^n$ (which is clearly
of Voronoi's first kind). This is a long standing open problem in the theory of lattices, which has
only recently been solved in certain very special cases~[6].
\\\textbf{Response:}
The final paragraph of Section~\zref{sec:discussion} has been modified to make the connection with the existing problem of determining whether a rectangular basis exists in a given lattice.

``A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.  It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.  Lattices of dimension larger than 3 need not be of Voronoi's first kind.  An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?  A related question is: is it possible to efficiently find an obtuse superbasis if it exists?  It is suspected that the answer to this second question is no.  An efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.  It is suspected that this simpler problem is difficult.''

@@

\item\textbf{Comment:} 
This paper makes a solid contribution to the theory of lattice problems. I think the results here are
well-motivated (I encourage the authors to include some of the motivation mentioned in the summary), and point to the possibility that there might be large and useful classes of lattices where
solving CVP is “easy”.
\\\textbf{Response:}
The following paragraph has been added to Section~\zref{sec:discussion} to suggest that similar fast algorithms for other classes of lattices might exist:

``Another interesting question is: are there subfamilies of Voronoi's first kind that admit even faster algorithms?  Both $A_n$ and $A_n^*$ are examples of this, but there might exist other subfamilies with algorithms faster than $O(n^4)$.  A related question is: can the techniques developed in this paper be applied to other families of lattices, i.e., beyond just those of Voronoi's first kind?''

\item\textbf{Comment:} 
On a technical level, I found the main convergence proof to be interesting, as I've rarely seen
combinatorial arguments successfully applied to lattice problems. On the other hand, while I've
convinced myself the authors have the ``right'' convergence proof, I found its presentation to be
rather awkward and unintuitive. I highly recommend that the authors rephrase the proof using
the language of Laplacians, which would make things more clear (see the additional comments
below for some technical suggestions).
\\\textbf{Response:}
We thank the reviewer for their comments.  We respond to the authors concerns below.

@@

We now respond to the reviewers numbered comments.

\begin{enumerate}

\item\textbf{Comment:} 
I believe Proposition 2.1 and 2.2 appear essentially verbatim in [8]. Same for Propositions
3.1 (which is essentially by definition) and 3.2, they appear in [9]. I don’t think it adds any
readibility/anything of value to reprove them here.
\\\textbf{Response:}

We agree that these propositions are well known/straightforward, but would prefer to keep them so that the paper is self contained. We do not know of references to specific proofs that we can give for these (unlike Theorem~4.1 for which a proof is directly given in \cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}).  The proof of Proposition 2.1 has been shortened in response to comment~\ref{com:rev2halfopen} of reviewer 2.  A result like Proposition 2.2 does appear informally in~\cite[Sec~5]{Shalvi_iterativeslicer_2009}.  As far as we are aware, the remarkable algorithm in~\cite{MicciancioVoulgaris_deterministic_jv_2013} does not rely upon bounding the number of lattice points inside a sphere (or any other volume).  Instead, the procedure operates recursively on scaled versions of the lattice.  The scaling is chosen to ensure that the target point always lies inside twice the Voronoi cell.

Motivating Propositions 3.1 and 3.2 is that the stopping criterion in our proceedure is slightly different from those in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013}.  The stopping criterions in~\cite{Shalvi_iterativeslicer_2009} and~\cite{MicciancioVoulgaris_deterministic_jv_2013} essentially involve a direct check that a given point, say $x$, is contained inside the Voronoi cell.  That is, one directly asserts that $2 x \cdot v \leq \|v\|^2$ for all relevant vectors $v$.  We cannot use this approach since it would amount to checking this inequality over as many as $2^{n+1}-2$ relevant vectors.  Instead we simply check that $\|y - x_K\| \leq \|y - x_{K+1}\|$.  That is, we check that the proceedure converges in the sense that 
\[
\|y - x_K\| = \|y - x_{K+1}\| = \|y - x_{K+2}\| = \dots.
\]
This is a straightforward, but important, difference and it motivates Propostions 3.1 and 3.2.

@@

\newcommand{\rng}{\operatorname{rng}}
\newcommand{\subrng}{\operatorname{subr}}
\newcommand{\decrng}{\operatorname{decrng}}

\item\textbf{Comment:} \label{com:uselaplaicesubr1}
Almost all the inequalities in the paper can be derived from the following simple observation.  For an $n$ vertex weighted graph $G$ with non-negative edge weights, and vectors, $\xbf,\ybf \in \reals^n$, if there exists a permutation $\pi$ of $[n]$ such that $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \leq \dots \leq \ybf_{\pi[n]}$ then
\[
\xbf^T L_G \ybf = \sum_{ \{i,j\} \in E[G] } \wbf_{ij} (\xbf_{i} - \xbf_{j}) (\ybf_{i} - \ybf_{j}) \geq 0.
\]
Furthermore, if $\xbf_{\pi[1]} \leq \dots \leq \xbf_{\pi[n]}$ and $\ybf_{\pi[1]} \geq \dots \geq \ybf_{\pi[n]}$ then the above inequality if reversed, i.e $\xbf^T L_G \ybf \leq 0$.

@@

I believe using the above will make many of your proofs more transparent.  For example, if you redefine $\subrng(\xbf)$ to return the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq 1$, then noice that if $S \neq \emptyset$ then there is a permutation that puts both $\xbf - 1_S$ and $1_S$ in non-decreasing order. In particular this implies that
\[
\xbf L_G \xbf = (\xbf - 1_S)^T L_G (\xbf - 1_S) + 2(\xbf - 1_S)^T L_G 1_S + 1_S^T L_G 1_S \geq (\xbf - 1_S)^T L_G (\xbf - 1_S),
\]
i.e. $\xbf$ can be made ``shorter'' by squeezing its components closer together.
\\\textbf{Response:}

The property decribed by the reviewer is indeed needed to give a proof of Theorem 4.1 (see our response to the next comment number~\ref{com:provethm4.1}).  However, this is not the property that is needed to prove that our algorithm converges in atmost $n$ steps.  To see this, let $\subrng_k(p)$ denote the largest set $S$ such that $\min_{i\in S} \xbf_i - \max_{j \notin S} \xbf_j \geq k$. Thus $\subrng_1$ is that suggested by the reviewer and $\subrng_2$ is that which is used in the paper.  Our proof works by considering a $v \in \ints^n$ that can be added to $\floor{z}$ so that $B(\floor{z} + v)$ is a closest lattice point to $y = Bz$.  Our Lemma~\zref{lem:roundzclose} shows that there exists such a $v$ with $\subrng_2(v) = \emptyset$ and so $\rng(v) \leq n$.  It is not true that there always exists such a $v$ with  $\subrng_1(v) = \emptyset$.

To consider this in more detail it is worth looking at the proof of Lemma~\zref{lem:decrngpreservesclosestpoints}.  The lemma shows that $B(\floor{z} + v)$ is a closest point to $y = Bz$ only if $B\big(\floor{z} + \decrng(v)\big)$ is.  A key part of the proof is the line
\[
1 + 2(\zeta_i-\zeta_j) - 2(v_i - v_j) < 1 + 2 - 4 = -1 < 0 \qquad \text{for $i \in S$ and $j \notin S$}
\]
where $\zeta_i = z_i - \floor{z_i} \in [0,1)$.  Since $\zeta_i-\zeta_j \in (-1,1)$ we require that the gap between $v_i - v_j$ to be greater than $\tfrac{3}{2}$.  Since $v \in \ints^{n+1}$ this means $v_i - v_j \geq 2$.  We cannot replace the current $\subrng_2$ with the suggested $\subrng_1$ and have this proof hold because then we would only have $v_i - v_j \geq 1$.

% Very early on (more than 2 years ago now!) when first looking at lattices of Voronoi's first kind, the first author believed that a single iteration of the proceedure would surfice.  That is, compute
% \[
% v = \arg\min_{t \in \{0,1\}^{n+1}}\| B(z - \floor{z} - t) \|^2
% \]
% and then $B(\floor{z} + v)$ would be a closest lattice point.  Numerical counterexamples dispelled this possibility.

@@

\item\textbf{Comment:} \label{com:provethm4.1}
Furthermore, one can use this to give a two line proof of Theorem 4.1.  This theorem I actually do think make sense to reprove here, because modifications of this proof are what drives the convergence bound. 
\\\textbf{Response:}

The property decribed by the reviewer is indeed needed to give a proof of Theorem 4.1.  To our knowledge, one also needs the property that the relevant vectors of a lattice are given by the shortest coset representatives of $\Lambda / 2\Lambda$.  This result is attributed to Voronoi (see for example \cite[Theorem~2]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases} and \cite[Theorem~3.2]{MicciancioVoulgaris_deterministic_jv_2013}).  This is used by Conway and Sloane in~\cite[Theorem~3]{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.

We believe that the proof, whilst not difficult, is ulitimately a distraction from our main results.  This would be different if properties needed for the proof, such as the function $\subrng_1$ and Selling's formula were also needed for proving the convergence of our algorithmm.  However, as we have stated in comments~\ref{com:uselaplaicesubr1}, Selling's formula is not directly needed, and $\subrng_2$, not $\subrng_1$ is the function required.

\item\textbf{Comment:} 
Lemma 5.1: there is a typo here.  The last identity should read
\[
\|B\pbf\|^2 - \|B(\pbf + 1_S - 1_T)\|^2 = \Phi(S,\pbf) + \Phi(\bar{T}, \pbf) + 2\sum_{i \in S}\sum_{j\in S} q_{ij}.
\]
Notice that factor 2 in front of the last term.  There are also two related typos on the top of page 12 where $\sum_{i \in S}\sum_{j\in S} q_{ij}$ (this appears twice on different lines) should be $2\sum_{i \in S}\sum_{j\in T} q_{ij}$.
\\\textbf{Response:}
This error has been fixed.

@@

\item\textbf{Comment:} 
Lemma 5.7: $h$ is ill-defined if $\rng(\vbf) = 0$.  You should explicitly state that $\rng(\vbf) \geq 1$.
\\\textbf{Response:}
Fixed.

\end{enumerate}

\section*{Reviewer 2}\label{sec:reviewer-2}

\begin{enumerate}
\item\textbf{Comment:}  
page 1, last line ``The closest lattice point problem is known to be NP-hard under certain condition ...'': I found this paragraph very confusing, and somehow inaccurate. What ``certain conditions''? I think a better way to describe the sitation regaring the NP-hardness of CVP in relation to this paper is the following. CVP is known to be NP-hard even to approximate within almost polynomial factor (see 35) when the lattice is arbitrary and not known in advance. Interestingly (see 35 and 17) there are families of lattices when CVP remain NP-hard even when given the best possible basis (or any hard-to-compute side information about the lattice). This version of the problem is called CVP with preprocessing (CVPP), and it is particularly relevant to this work, where it is shown that CVPP can be solved exactly and in polynomial time for the class of lattices of Voronoi first kind, when given as auxiliary information a special basis (namely, an obtuse superbasis.) 
\\\textbf{Response:}

@@

The interested reader could read these references in the order given. %, i.e., first Chapter 4 of~\cite{SPLAG} and~\cite{Conway1982FastQuantDec} for $D_n$ and $D_n^*$ and older algorithms for $A_n$ and $A_n^*$.  Then~\cite{McKilliam2008} for a fast algorithm for $A_n^*$, then~\cite{McKilliam2009CoxeterLattices} for fast algorithms for all of the Coxeter lattices that includes $A_n$ and $A_n^*$, $E_8$, etc.  Technically, the algorithm from~\cite{McKilliam2009CoxeterLattices} does include that from~\cite{McKilliam2008}.  However~\cite{McKilliam2008} is simpler and shorter to read first.

\item\textbf{Comment:}  
When saying ``if the obtuse superbasis is known'', you may want to cast this requirement in the context of CVPP. (The obtuse superbasis is the advice provided by the preprocessing.) 
\\\textbf{Response:}

\item\textbf{Comment:}  
page 2,bottom, displayed equation "$v \cdot x \leq x \cdot x$". Maybe you mean strict inequality.
\\\textbf{Response:}
Our definition includes what Conway and Sloane call the `lax' relevant vectors.  To make this clear the following sentences have been added below the definition of the relevant vectors in Section~\zref{sec:voron-cells-relev}:

``This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered.'' %It will be of benefit to include all of the strict and lax relevant vectors here.''

\item\textbf{Comment:}  
page 3, and maybe elsewhere ``short vector in a lattice'' should probably be ``shortest nonzero vector in a lattice''. 
\\\textbf{Response:}
We have instead introduced the term `short vector' in the introduction.  The following sentence is included in the first paragraph of the introduction: ``This is called the \emph{shortest vector problem} and a solution is called a~\emph{short vector}.''

@@

\item\textbf{Comment:}  
page 3, definition of ceiling(r) as smallest integer strctly larger than 3 seems unconventional. The standard definition is not strict. Write floor(r)+1 instead, which is consistent with the definition of floor(r) given later in the paper. 
\\\textbf{Response:}
Done.

### 2014-07-10 reviews/response.tex

\item\textbf{Comment:}  \label{com:rev2halfopen}
page 3, proof of Prop. 2.1. Your definition of ``half-open voronoi cell'' is not a proper definition, and also seems incorrect. E.g., take the integer lattice $\ints^n$, then only one of the corners (rather than half) of the corners of the Voronoi cell should be included if you want a tiling. 
\\\textbf{Response:}
The reviewer is correct.  We only require that some modification of the boundary of the closed Voronoi cell results in a tiling.  The proof of Proposition~\zref{eq:latticepointsinvorcvell} has been modified accordingly.

### 2014-07-10 paper.tex
minor
added minor

Equivalently, the Voronoi cell can be defined as the intersection of the half spaces 
\begin{align*}
H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
\end{align*}
for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.   %Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.  A lattice point in $\calR$ is called \term{relevant}. The following remark follow trivially from the preceding discussion.
The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
\[
\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$}.
\]
This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered. It will be of benefit to include all of the strict and lax relevant vectors here.  Denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
\[
%\begin{equation}\label{eq:defvortermsrelvecs}
\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
%\end{equation}
\]
The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
\begin{equation}\label{eq:relvectnearpointieq}
\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
\end{equation}

@@

A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.  It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.  Lattices of dimension larger than 3 need not be of Voronoi's first kind.  An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?  A related question is: is it possible to efficiently find a obtuse superbasis if it exists?  It is suspected that the answer to both of these questions is no.  An efficient solution to either question would immediately yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.

>>>>

Equivalently, the Voronoi cell can be defined as the intersection of the half spaces 
\begin{align*}
H_{v} &= \{x \in \reals^n \mid \|x\| \leq \|x - v\| \} \\
&= \{x \in \reals^n \mid \dotprod{x}{v} \leq \tfrac{1}{2}\dotprod{v}{v} \}
\end{align*}
for all $v \in \Lambda \backslash  \{\zerobf\}$.  %We denote by $x^\prime$ the transpose of the vector $x$ and so $\dotprod{x}{v}$ is the inner product between column vectors $x$ and $v$.  
We denote by $\dotprod{x}{v}$ the inner product between vectors $x$ and $v$.
It is not necessary to consider all $v \in \Lambda \backslash  \{\zerobf\}$ to define the Voronoi cell.   %Figure \ref{lattices:fig:vorregion} depicts the relevant vectors of the lattice with generator given by~\eqref{lattices:eq:genB}.  A lattice point in $\calR$ is called \term{relevant}. The following remark follow trivially from the preceding discussion.
The \emph{relevant vectors} are those lattice points $v \in \Lambda \backslash  \{\zerobf\}$ for which  
\[
\dotprod{v}{x} \leq \dotprod{x}{x} \qquad \text{for all $x \in \Lambda$}.
\]
This definition includes what Conway and Sloane call the `lax' relevant vectors~\cite{ConwaySloane1992_voronoi_lattice_3d_obtuse_superbases}.  If the definition above is replaced by one with strict inequality then only the `strict' relevant vectors would be considered. %It will be of benefit to include all of the strict and lax relevant vectors here.  
Denote by $\relevant(\Lambda)$ the set of relevant vectors of the lattice $\Lambda$.  The Voronoi cell is the intersection of the halfspaces corresponding with the relevant vectors, that is, 
\[
%\begin{equation}\label{eq:defvortermsrelvecs}
\vor(\Lambda) = \cap_{v\in\relevant(\Lambda)}{H_{v}}.
%\end{equation}
\]
The closest lattice point problem and the Voronoi cell are related in that $x\in\Lambda$ is a closest lattice point to $y$ if and only $y - x \in \vor(\Lambda)$, that is, if and only if
\begin{equation}\label{eq:relvectnearpointieq}
\dotprod{(y - x)}{v} \leq \tfrac{1}{2} \dotprod{v}{v} \qquad \text{for all $v \in \relevant(\Lambda)$}.  
\end{equation}

@@

A final remark is that our algorithm assumes that the obtuse superbasis is known in advance.  It is known that all lattices of dimension less than or equal to 3 are of Voronoi's first kind and an algorithm exists to recover the obtuse superbasis in this case~\cite{SPLAG}.  Lattices of dimension larger than 3 need not be of Voronoi's first kind.  An interesting question is: given a lattice, is it possible to efficiently decide whether it is of Voronoi's first kind?  A related question is: is it possible to efficiently find an obtuse superbasis if it exists?  It is suspected that the answer to this second question is no.  An efficient solution would yield a solution to a known problem, that of determining whether a lattice is rectangular (has a basis consisting of pairwise orthogonal vectors) given an arbitrary basis~\cite{Lenstra_Silverberg_revisting_gentra_szydlo_2014}.  It is suspected that this simpler problem is difficult.
